---
title: "THRESHOLD DETERMINATION"
author: "Gorka Campandegui García"
date: "`r Sys.Date()`"
output: pdf_document
---

Libraries to be used:

```{r, message = FALSE, warning=FALSE}
library(ggplot2)
library(POT)
library(evd)
library(splines)
library(pander)
library(tea)
library(dplyr)
library(multcomp)
```

\tableofcontents

\newpage

# INTRODUCTION

```{r}
g_xi_positive <- function(y, sigma, xi) (1/sigma) * (1 + xi*y/sigma) ^ (- 1 - 1/xi)
g_xi_zero <- function(y, sigma) (1/sigma) * exp(-y/sigma)
g_xi_negative <- function(y, sigma, xi) {
  ((1/sigma) * (1 + xi*y/sigma) ^ (- 1 - 1/xi)) * (y< -sigma/xi)}
y <- seq(0, 5, 0.05)
sigma <- 1
xi1 <- 0.25
xi2 <- -0.25
data1 <- data.frame(y, g_xi_positive(y, sigma, xi1), g_xi_negative(y, sigma, xi2),
                    g_xi_zero(y, sigma))
colnames(data1) <- c("y", "g_pos", "g_neg", "g_zero")
ggplot2::ggplot(data = data1, aes(x = y, y = g_pos)) + theme_minimal() +
  geom_line(aes(color = "deepskyblue"), show.legend = TRUE) +
  geom_line(aes(x = y, y = g_neg, color = "deeppink"), show.legend = TRUE) +
  geom_line(aes(x = y, y = g_zero, color = "chartreuse3"), show.legend = TRUE) +
  ggtitle("Generalized Pareto density function") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("y") +
  ylab("g(y)") +
  scale_color_manual(name = "Cases",
                     values = c("deepskyblue", "deeppink", "chartreuse3"),
                     labels = c("Pareto type", "Upper endpoint type", "Exponential type"))
```

\newpage

# METHODS FOR THRESHOLD SELECTION

## Mean Residual Life Plot

```{r, warning=FALSE, comment=""}
# FUNCTION FOR PLOTTING THE MRLP

mrlp <- function(data, u_min=0.1, n0=10){
  
  # Transformation of the data
  d <- as.matrix(data)
  
  # Compute the independent peaks over threshold
  pot <- evd::clusters(data = d, u = u_min, cmax = TRUE)
  pot <- sort(pot)
  
  # Mean excess and quasi-standard deviation computation
  n <- length(pot)
  e <- c()
  s <- c()
  for (i in 1:(n-n0)){
    excesses <- pot[ pot >= pot[i]] - pot[i]
    e[i] <- mean(excesses)
    }
  u <- pot[1:(n-n0)]
  
  # Plotting
  df <- data.frame(u, e)
  ggplot2::ggplot(data = df, aes(x = u, y = e)) + geom_line(color='grey') +
    geom_point(color = "blue2", size = 0.6) + theme_minimal() +
    labs(x = "Threshold", y = "Mean of the excesses",
       title = "Mean Residual Life Plot") +
    theme(plot.title = element_text(hjust = 0.5))
}
```

\newpage

## WMSE criteria by Langousis

```{r}
wmse_langousis <- function(data, u_min, n0=10, log_scale=FALSE, plot=FALSE){
  
  # Transformation of the data
  d <- as.matrix(data)  
  
  # Peaks over threshold selection
  pot <- evd::clusters(data = d, u = u_min, cmax = TRUE)
  pot <- sort(pot)
  
  # Mean excess, quasi-standard deviation and weights computation
  n <- length(pot)
  e <- c()
  s <- c()
  
  for (i in 1:(n-n0)){
    excesses <- pot[ pot >= pot[i]] - pot[i]
    e[i] <- mean(excesses)
    s[i] <- sd(excesses)
    }
  
  u <- pot[1:(n-n0)]
  j <- seq(1,n-n0,1)
  w <- (n-j)/(s^2)
  
  # WMSE
  wmse <- c()
  for (i in 1:(n-2*n0)){
    ee <- e[i:(n-2*n0)]
    uu <- u[i:(n-2*n0)]
    ww <- w[i:(n-2*n0)]
    model <- lm(ee ~ uu, weights = ww)
    residuals <- resid(model)
    wmse_i <- mean(ww*residuals^2)
    wmse[i] <- wmse_i
  }
  
  # Threshold
  i <- 1
  while(wmse[i+1]<wmse[i]){
    i <- i+1
  }
  threshold <- u[i]

  # Plotting
  if(plot==TRUE){  
    
  u2n0 <- u[1:(n-2*n0)]
  df <- data.frame(u2n0, wmse)
  
  if(log_scale == TRUE){
    # Plot in logarithmic scale
    pl <- ggplot2::ggplot(data = df, aes(x = u2n0, y = wmse)) + geom_line(color='grey') +
      scale_y_continuous(trans='log10') +
      geom_point(color = "blue2", size = 0.6) + theme_minimal() +
      geom_vline(xintercept = threshold, color="red")+
      labs(x = "Threshold", y = "log(WMSE)",
       title = "Threshold selection based on WMSE criteria by Langousis (2016)") +
      theme(plot.title = element_text(hjust = 0.5))
  }else{
    # Plot in original scale
    pl <- ggplot2::ggplot(data = df, aes(x = u2n0, y = wmse)) + geom_line(color='grey') +
      geom_point(color = "blue2", size = 0.6) + theme_minimal() +
      geom_vline(xintercept = threshold, color="red")+
      labs(x = "Threshold", y = "WMSE",
       title = "Threshold selection based on WMSE criteria by Langousis (2016)") +
      theme(plot.title = element_text(hjust = 0.5))
  }
    result <- list(plot=pl, threshold=as.numeric(threshold))
  }else{
    result <- list(threshold=as.numeric(threshold))
  }
  
  return(result)
  
  }
```

\newpage

## Goodness of Fit methods

### Cramer-Von Mises

```{r}
threshold_CVM <- function(data, alpha=0.05){
  
    # Reduce the data to the upper 90%
  data <- data[data>quantile(data,0.9)]
  
  # Sort and convert data to matrix
  data <- as.matrix(sort(data))
  
  # Set the initial values
  p_value <- 0
  i <- 1
  
  # While a GPD is rejected, do the following
  while(p_value < alpha){
    
    # Define the current threshold
    u <- data[i]
    
    # Compute the excesses
    excesses <- data[data>u]
    
    # Stop if there are less than 10 excesses
    if (length(excesses) < 10) {
      break
      }
    
    # Fitting of the GPD
    parameters <- tea::gpdFit(excesses, threshold=u)$par.ests
    scale <- as.numeric(parameters[1])
    shape <- as.numeric(parameters[2])
    
    # Cramer - Von Mises test
    PGPD <- function(q){tea::pgpd(scale=scale, shape=shape, loc=u, q=q)}
    n <- length(excesses)
    k <- 1:n
    prob <- PGPD(excesses)
    prob <- prob[prob<1]
    qnor <- qnorm(sort(prob))
    pnor <- pnorm((qnor-mean(qnor))/sd(qnor))
    w <- round((sum((pnor-(2*k-1)/(2*n))^2)+1/(12*n))*(1+0.5/n),4)
    if (w<0.0275){
      p_value <- 1-exp(-13.953+775.5*w-12542.61*w^2)
    }
    if (0.0275 <= w & w < 0.051){
      p_value <- 1-exp(-5.903+179.546*w-1515.29*w^2)
    }
    if (0.051 <= w & w < 0.092){
      p_value <- exp(0.886 - 31.62 * w + 10.897 * w^2)
    }
    if (0.092 <= w & w < 1.1){
      p_value <- exp(1.111 - 34.242 * w + 12.832 * w^2)
    }
    if (w >=1.1){
      p_value <- 7.37 * 10^(-10)
    }

    # Go to next value
    i <- i+1
  }
  return(threshold=u)
}

```    
    
\newpage
    
### Anderson - Darling

```{r}
threshold_AD <- function(data, alpha){
  
  # Reduce the data to the upper 90%
  data <- data[data>quantile(data,0.9)]
  
  # Sort and convert data to matrix
  data <- as.matrix(sort(data))
  
  # Set the initial values
  p_value <- 0
  i <- 1
  
  # While a GPD is rejected, do the following
  while(p_value < alpha){
    
    # Define the current threshold
    u <- data[i]
    
    # Compute the excesses
    excesses <- data[data>u]
    
    # Stop if there are less than 10 excesses
    if (length(excesses) < 10) {
      break
      }
    
    # Fitting of the GPD
    parameters <- tea::gpdFit(excesses, threshold=u)$par.ests
    scale <- as.numeric(parameters[1])
    shape <- as.numeric(parameters[2])
    
    # Anderson - Darling test
    PGPD <- function(q){tea::pgpd(scale=scale, shape=shape, loc=u, q=q)}
    n <- length(excesses)
    k <- 1:n
    prob <- PGPD(excesses)
    prob <- prob[prob<1]
    qnor <- qnorm(sort(prob))
    pnor <- pnorm((qnor-mean(qnor))/sd(qnor))
    A <- (-n-sum((2*k-1)*log(pnor)+(2*n+1-2*k)*log(1-pnor))/n)*(1+0.75/n+2.25/n^2)
    A <- round((1 + 0.75/n + 2.25/n^2) * A,4)
    if (A < 0.2){
      p_value <- 1 - exp(-13.436+101.14*A-223.73*A^2)
    }
    if (0.2 <= A & A < 0.34){
      p_value <- 1 - exp(-8.318 + 42.796 * A - 59.938 * A^2);
    }
    if (A <= 0.34 & A < 0.6){
      p_value <- exp(0.9177 - 4.279 * A - 1.38 * A^2)
    }
    if (A <= 0.6 & A < 10){
      p_value <- exp(1.2937 - 5.709 * A + 0.0186 * A^2)
    }
    if (A >= 10){
      p_value <- 7.37*10^(-10)
    }
  
    # Go to next value
    i <- i+1
  }
  return(threshold=u)
}
```

\newpage

## Threshold selection based on studentized residuals

```{r}
threshold <- function(data, u_min, n0=10, plot=FALSE, deg=25, alpha=0.05){

  # Transformation of the data
  d <- as.matrix(data)
  
  # Peaks over threshold selection
  pot <- evd::clusters(data = d, u = u_min, cmax = TRUE)
  pot <- sort(pot)
  
  # Loop that repeats until condition is satisfied
  condition <- FALSE
  
  # We define some auxiliary variables so that the plot is made with the result of the
  # previous iteration
  u <- 0
  u0 <- 0
  zeros_indices <- 0
  studentized_residuals <- 0
  predicted_sr <- 0
  sr0 <- 0
  while (condition==FALSE){
    threshold <- u_min
    u_ <- u
    u0_ <- u0
    zeros_indices_ <- zeros_indices
    studentized_residuals_ <- studentized_residuals
    predicted_sr_ <- predicted_sr
    sr0_ <- sr0
    pot <- pot[pot>=u_min]
  
  # Mean excess and quasi-standard deviation computation
  n <- length(pot)
  e <- c()
  s <- c()
  for (i in 1:(n-n0)){
    excesses <- pot[pot >= pot[i]] - pot[i]
    e[i] <- mean(excesses)
    s[i] <- sd(excesses)
    }
  u <- pot[1:(n-n0)]
  j <- seq(1,n-n0,1)
  w <- (n-j)/(s^2)
  
  # Define the linear model and compute internally studentized residuals
  model <- lm(e ~ u, weights = w)
  studentized_residuals <- rstandard(model)
  
  # Compute the objective function to see whether there is bad data
  res <- resid(model)
  obj <- sum(w*res^2)
  condition <- obj < qchisq(1-alpha, n-n0-2)
  
  # Fit a spline
  spline <- lm(studentized_residuals ~ bs(u, degree = deg))
  predicted_sr <- predict(spline)
  
  # Indices of the zeros of the spline
  zeros_indices <- c()
   for (i in 1:(length(u)-1)){
     if (sign(predicted_sr[i+1]) != sign(predicted_sr[i])){
       zeros_indices <- c(zeros_indices, i)
      }
    }
  
  # Give the threshold
  ind1 <- zeros_indices[1]
  ind2 <- zeros_indices[2]
  threshold_index <- which(abs(studentized_residuals[ind1:ind2]) == max(abs(studentized_residuals[ind1:ind2]))) + ind1
  u_min <- as.numeric(u[threshold_index])
  
  # The while loop finishes here
}
  
  # Plot
  if(plot == TRUE){
    # Plot 1: how the threshold has been determined
    u0_ <- u_[zeros_indices_]
    sr0_ <- predicted_sr_[zeros_indices_]
    df0 <- data.frame(u_, studentized_residuals_, predicted_sr_)
    df1 <- data.frame(u0_, sr0_)
    pl1 <- ggplot2::ggplot(data = df0, aes(x = u_, y = studentized_residuals_)) +
      theme_minimal() +
      geom_point(color = "blue2", size = 0.6) +
      geom_line(data = df0, aes(x=u_, y=predicted_sr_, group=1), color = "red") +
      geom_vline(data = df1, aes(xintercept = u0_),
             linetype = "dashed", color = "black", alpha = 0.5) +
      geom_vline(data = df1, aes(xintercept = threshold),
             linetype = "solid", color = "lightgreen") +
      labs(x="Threshold", y="Internally studentized residuals")
    
    # Plot 2: studentized residuals of the regression of the excesses
    u0 <- u[zeros_indices]
    sr0 <- predicted_sr[zeros_indices]
    df0 <- data.frame(u, studentized_residuals, predicted_sr)
    df1 <- data.frame(u0, sr0)
    pl2 <- ggplot2::ggplot(data = df0, aes(x = u, y = studentized_residuals)) +
      geom_point(color = "blue2", size = 0.6) + theme_minimal() +
      geom_vline(data = df1, aes(xintercept = threshold),
                 linetype = "solid", color = "lightgreen") +
      labs(x="Threshold", y="Internally studentized residuals")
    
    if(u_[1]==0){
      result <- list(plot = pl2, threshold = threshold)
    }else{
    result <- list(plot1 = pl1, plot2 = pl2, threshold = threshold)
    }
  }else{
      result <- list(threshold=threshold)
  }
  return(result)
  }
```


\newpage

# APPLICATION OF THE METHODS

## NOAA data

Loading the data and brief descriptive analysis:

```{r, comment=""}
# We load the data
data <- read.csv("PRCP_ASN00021043.csv", sep=",")

# Dimensions of the dataset
dim(data)

# We display the number of missing values
sum(is.na(data))

# We assign 0 to NAs
data[is.na(data)] <- 0

# We check that there is no missing data now
sum(is.na(data))

# We divide the data by 10
data["PRCP"] <- data["PRCP"]/10

# Number of 0s
sum(data['PRCP']==0)

# Proportion of 0s
sum(data['PRCP']==0)/46132
```

Plot the data:

```{r, fig.height=3, warning=FALSE}
# We plot the data
ggplot2::ggplot(data=data, aes(x=DATE, y=PRCP)) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  geom_segment(aes(x = DATE, xend = DATE, y = 0, yend = PRCP), color = "gray", size = 0.5) +
  geom_point(color = "blue2", size = 0.6)

# Zoom in year 2000
data0 <- data[41506:41871,]
ggplot2::ggplot(data=data0, aes(x=DATE, y=PRCP)) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  geom_segment(aes(x = DATE, xend = DATE, y = 0, yend = PRCP), color = "gray", size = 0.5) +
  geom_point(color = "blue2", size = 0.6)
```

Compute the threshold:

```{r}
# Threshold selection techniques
mrlp(data[,'PRCP'])
wmse_langousis(data[,'PRCP'], u_min=0.1, plot=TRUE)
threshold(data[,'PRCP'], plot=TRUE, u_min=0.1)
threshold_AD(data=data[,'PRCP'], alpha=0.005)
threshold_CVM(data=data[,'PRCP'], alpha=0.005)
```

\newpage

## Simulated data

```{r}
# Set the seed
set.seed(4545454)

# Establish the parameters
mu <- 0
sigma_2 <- 4
beta <- 0.975
sigma <- 1
xi <- 0.15
n <- 10^4
u <- qnorm(beta, mean=mu, sd=sqrt(sigma_2))

# Generate the data
s <- rnorm(n, mean=mu, sd=sqrt(sigma_2))
for (i in 1:length(s)){
  if (s[i]>u){
    s[i] <- rgpd(1, loc = u, scale = sigma, shape = xi)
  }
  if (s[i]<0){
    s[i] <- 0
  }
}

# Show the theshold
print(paste("Threshold=",u))

# Plot the generated data
data2 <- data.frame(1:n,s)
colnames(data2) <- c("Time", "Value")
ggplot2::ggplot(data=data2, aes(x=Time, y=Value)) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  geom_segment(aes(x = Time, xend = Time, y = 0, yend = Value), color = "gray", linewidth = 0.5) +
  geom_point(color = "blue2", size = 0.6) + theme_minimal() +
  geom_hline(data = data2, aes(yintercept = u),
                 linetype = "solid", color = "red")

# Plot the density
ggplot2::ggplot(data=data2, aes(x=Value)) +
  geom_density(color = "blue2", fill="blue2", alpha=0.2) + theme_minimal() +
  geom_vline(data = data2, aes(xintercept = u), color = "red") +
  annotate("text", x = u+0.8, y = 1, label = paste("Threshold=",u),
           color = "red", size = 4) + labs(y="Density")
  
```

Find the threshold with several methods:

```{r}
# Threshold selection techniques
mrlp(s)
wmse_langousis(s, u_min=0.1, plot=TRUE)
threshold(s, plot=TRUE, alpha=0.05, u_min=0.1)
threshold_AD(data=s, alpha=0.05)
threshold_CVM(data=s, alpha=0.05)
```

\newpage

## Data from Igeldo

```{r}
# Load the data
igeldo <- read.csv("igeldoHour.csv")
igeldo.prec <- as.vector(igeldo$Prec[!is.na(igeldo$Prec)])

# Threshold selection techniques
mrlp(igeldo.prec)
wmse_langousis(igeldo.prec, u_min=0.1, plot=TRUE)
threshold(igeldo.prec, alpha=0.05, u_min=0.1, plot=TRUE)
# threshold_AD(data=igeldo.prec, alpha=0.05)
# threshold_CVM(data=igeldo.prec, alpha=0.05)
```


\newpage

# ANALYSIS AND COMPARISONS OF THE METHODS

## Simulation of data, threshold computation & storing the information

```{r, warning=FALSE, eval=FALSE}
# Establish the parameters
mu <- 0
sigma_2 <- 1

# Establish the parameter grid
betagrid <- c(0.95,0.98,0.99,0.995)
sigmagrid <- c(0.1,0.5,5,10)
xigrid <- c(-0.15,0,0.15)
ngrid <- c(2000,10^4)
alphagrid <- c(0.1,0.05,0.005,0.001)

# Number of experiments
N <- 4*4*3*2*4*4

# Create a dataframe
anova.df <- data.frame(
  "difference" = rep(0,N),
  "method" = rep("",N),
  "beta" = rep(0,N),
  "sigma" = rep(0,N),
  "xi" = rep(0,N),
  "n" = rep(0,N),
  "alpha" = rep(0,N),
  "time" = rep(0,N),
  "hatsigma" = rep(0,N),
  "hatxi" = rep(0,N)
)

# Generate the data and compute the threshold
i <- 1
for (beta in betagrid){
  for (sigma in sigmagrid){
    for (xi in xigrid){
      for (n in ngrid){
        for (alpha in alphagrid){
          
          # Set the threshold
          u <- qnorm(beta, mean=mu, sd=sqrt(sigma_2))

          # Simulate the data
          s <- rnorm(n, mean=mu, sd=sqrt(sigma_2))
          for (k in 1:length(s)){
            if (s[k]>u){
              s[k] <- rgpd(1, loc = u, scale = sigma, shape = xi)
            }
            if (s[k]<0){
              s[k] <- 0
            }
          }
          
          # Compute the threshold
          
          t0 <- Sys.time()
          tL <- wmse_langousis(s, u_min = 0.1)$threshold
          t1 <- Sys.time()
          dif <- as.numeric(t1-t0)
          parameters <- tea::gpdFit(s[s>tL], threshold=tL)$par.ests
          hatscale <- as.numeric(parameters[1])
          hatshape <- as.numeric(parameters[2])
          anova.df[i,] <- c(tL-u,"Langousis",beta,sigma,xi,n,alpha,dif,hatscale,hatshape)
          print(paste(i, " of ", N, "  (Langousis)"))
          
          t0 <- Sys.time()
          tR <- threshold(s, alpha=alpha, u_min=0.1)$threshold
          t1 <- Sys.time()
          dif <- as.numeric(t1-t0)
          parameters <- tea::gpdFit(s[s>tR], threshold=tR)$par.ests
          hatscale <- as.numeric(parameters[1])
          hatshape <- as.numeric(parameters[2])
          anova.df[i+1,] <- c(tR-u,"Residuals",beta,sigma,xi,n,alpha,dif,hatscale,hatshape)
          print(paste(i+1, " of ", N, "  (Residuals)"))
          
          t0 <- Sys.time()
          tAD <- threshold_AD(s, alpha=alpha)
          t1 <- Sys.time()
          dif <- as.numeric(t1-t0)
          parameters <- tea::gpdFit(s[s>tAD], threshold=tAD)$par.ests
          hatscale <- as.numeric(parameters[1])
          hatshape <- as.numeric(parameters[2])
          anova.df[i+2,] <- c(tAD-u,"AD",beta,sigma,xi,n,alpha,dif,hatscale,hatshape)
          print(paste(i+2, " of ", N, "  (AD)"))
          
          t0 <- Sys.time()
          tCVM <- threshold_CVM(s, alpha=alpha)
          t1 <- Sys.time()
          dif <- as.numeric(t1-t0)
          parameters <- tea::gpdFit(s[s>tCVM], threshold=tCVM)$par.ests
          hatscale <- as.numeric(parameters[1])
          hatshape <- as.numeric(parameters[2])
          anova.df[i+3,] <- c(tCVM-u,"CVM",beta,sigma,xi,n,alpha,dif,hatscale,hatshape)
          print(paste(i+3, " of ", N, "  (CVM)"))
          
          # Change the index
          i <- i+4
          
        }
      }
    }
  }
}

```

The results are stored in the file named *anova.csv*.

## Analysis of the differences among methods

```{r}
# Read the data
df <- read.csv("anova.csv")
```

### GENERAL ASPECTS

#### Threshold detection // Sample size

```{r, warning=FALSE}
df$n <- as.factor(df$n)

ggplot(df, aes(x = difference, fill = n)) +
  geom_density(alpha = 0.3) + 
  xlim(-3, 3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Density")

```

```{r}
dif_short <- df$difference[df$n == "2000"]
dif_long <- df$difference[df$n == "10000"]
mean(dif_long)
mean(dif_short)
mean(dif_long) - mean(dif_short)
wilcox.test(dif_long, dif_short)
```

#### Threshold detection // Sigma

```{r, warning=FALSE}
df$sigma <- as.factor(df$sigma)

ggplot2::ggplot(df, aes(x=difference, fill=sigma)) +
  geom_density(alpha=0.3) + xlim(-3,3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Density")
```

#### Threshold detection // Xi

```{r, warning=FALSE}
df$xi <- as.factor(df$xi)

ggplot2::ggplot(df, aes(x=difference, fill=xi)) +
  geom_density(alpha=0.3) + xlim(-3,3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Density")
```

#### Threshold detection // Beta

```{r, warning=FALSE}
df$beta <- as.factor(df$beta)

ggplot2::ggplot(df, aes(x=difference, fill=beta)) +
  geom_density(alpha=0.3) + xlim(-3,3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Density")
```

#### Threshold detection // Alpha

```{r, warning=FALSE}
df$alpha <- as.factor(df$alpha)

ggplot2::ggplot(df, aes(x=difference, fill=alpha)) +
  geom_density(alpha=0.3) + xlim(-3,3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Density")
```

### DIFFERENCE // METHOD

```{r, warning=FALSE, message=FALSE}
# Convert to factor
df$method <- as.factor(df$method)
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=difference, fill=method)) +
  geom_density(alpha=0.3) + xlim(-3,2) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Density")
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=difference, fill=method)) +
  geom_boxplot() + xlim(-3,5) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab(expression(Difference~(hat(u) - u))) + 
  ylab("Method") +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

Check normality of the difference by groups:

```{r}
dif_res <- df$difference[df$method == "Residuals"]
dif_lan <- df$difference[df$method == "Langousis"]
dif_and <- df$difference[df$method == "AD"]
dif_cvm <- df$difference[df$method == "CVM"]
shapiro.test(dif_res)$p.value
shapiro.test(dif_lan)$p.value
shapiro.test(dif_and)$p.value
shapiro.test(dif_cvm)$p.value
```

Clearly not normal. We use Kruskal-Wallis test as an alternative to ANOVA for non Gaussian data:

```{r}
kruskal.test(difference ~ method, data=df)$p.value
```

Clearly there are differences among groups. Wilcoxon-Mann-Whitney tests:

```{r}
wilcox.test(dif_res, dif_lan)$p.value # DIFFERENT
wilcox.test(dif_res, dif_and)$p.value # ?
wilcox.test(dif_res, dif_cvm)$p.value # EQUAL
wilcox.test(dif_lan, dif_and)$p.value # DIFFERENT
wilcox.test(dif_lan, dif_cvm)$p.value # DIFFERENT
wilcox.test(dif_and, dif_cvm)$p.value # DIFFERENT
```

### TIME // METHOD

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=time, fill=method)) +
  geom_density(alpha=0.3) + xlim(0,18) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab("Time (s)") + 
  ylab("Density")
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=time, fill=method)) +
  geom_boxplot() + xlim(0,18) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab("Time (s)") + 
  ylab("Method")
```

```{r}
time_res <- df$time[df$method == "Residuals"]
time_lan <- df$time[df$method == "Langousis"]
time_and <- df$time[df$method == "AD"]
time_cvm <- df$time[df$method == "CVM"]
shapiro.test(time_res)$p.value
shapiro.test(time_lan)$p.value
shapiro.test(time_and)$p.value
shapiro.test(time_cvm)$p.value
```

Not normal.

```{r}
kruskal.test(time ~ method, data=df)$p.value
```

Significant differences.

```{r}
wilcox.test(time_res, time_lan)$p.value # DIFFERENT
wilcox.test(time_res, time_and)$p.value # DIFFERENT
wilcox.test(time_res, time_cvm)$p.value # DIFFERENT
wilcox.test(time_lan, time_and)$p.value # ?
wilcox.test(time_lan, time_cvm)$p.value # ?
wilcox.test(time_and, time_cvm)$p.value # DIFFERENT
```

### TIME // SAMPLE SIZE

```{r}
time_short <- df$time[df$n == "2000"]
time_long <- df$time[df$n == "10000"]
mean(time_long)
mean(time_short)
mean(time_long) - mean(time_short)
wilcox.test(time_long, time_short)
```

### PARAMETER ESTIMATION

#### Sigma

```{r}
df <- read.csv("anova.csv")
df$rel_sigma <- (df$hatsigma - df$sigma) / df$sigma
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=rel_sigma, fill=method)) +
  geom_density(alpha=0.3) + xlim(-3,13) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab("Relative error when estimating sigma") + 
  ylab("Density")
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=rel_sigma, fill=method)) +
  geom_boxplot() + xlim(-3,18) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab("Relative error when estimating sigma") + 
  ylab("Method") +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

#### Xi

```{r}
df$error_xi <- df$hatxi - df$xi
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=error_xi, fill=method)) +
  geom_density(alpha=0.3) + xlim(-3,3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab("Absolute error when estimating xi") + 
  ylab("Method")
```

```{r, warning=FALSE}
ggplot2::ggplot(df, aes(x=error_xi, fill=method)) +
  geom_boxplot() + xlim(-3,3) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() + 
  xlab("Absolute error when estimating xi") + 
  ylab("Method") +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```
